---
title: "Aviation Bio-Fuel Stock Sourcing"
output: html_notebook
---

# Problem Statement

A company that produces aircraft biofuel is planning a new product
called FC (Fuel-Corn). The table below shows the total quarterly demand in
tons (t) for the coming years as communicated by their customers.

FC demand (T)

| Q1             | Q2           | Q3           | Q4          |
| :------------- | :----------: | -----------: |-----------: |
| 1,200    | 1,100    | 1,300   |  1,000 |

Fuel corn is made of a combination of Fuel and Corn. The supply schedule dictates that we purchase these raw materials on a bimonthly schedule. The costs per ton of Fuel and Corn for
every two month period in the coming year is listed below.

|           | B1        | B2      |    B3    | B4   |  B5  |  B6  |
| :-------- | :-------: | ------: |--------: | ---: |----: |----: |
|Fuel (\$/t)| 2         |2.5      | 2        |  1   |  1.5 |  3   |
Corn (\$/t) | 1.5       | 1       |   2     |  1|  2|   2.5|


FC composition is obtained by mixing 35% of Fuel and 65% of Corn.
The life of Fuel is of four consecutive months and the life of Corn, six
(i.e., if we buy Fuel in early January, we cannot use it in early May).
We just buy Fuel and Corn at the beginning of each two-month period
and make the deliveries of FC at the beginning of each quarter. For
simplicity, we assume that one can buy, mix and sell the same day.
In addition, the plant manager has told us that in any two-month period, we cannot buy more Fuel than triple of Corn.

Develop a model to minimize the cost of production.

# Problem Setup
```{r}
library(lpSolve)

#fuel_life <- 4
#corn_life <-6
#buy_cycle <- 2
deliveries <- c(1200, 1100, 1300, 1000)
required_fuel <- deliveries*0.35
print(required_fuel)
required_corn <- deliveries*0.65
print(required_corn)
#delivery_dates <- c(1,4,7,10)
#purchase_dates <- c(1,3,5,7,9,11)
fuel_costs <- c(2,2.5,2,1,1.5,3)
corn_costs <- c(1.5,1,2,1,2,2.5)
```

```{r}
constr <- matrix(c(first_delivery_fuel, first_delivery_corn), ncol=12, byrow=TRUE)
constr.dir <- c(">=", ">=")
rhs <- c(required_fuel[1], required_corn[1])

obj.fun <- c(fuel_costs, corn_costs)

#Preview problem definition
print(obj.fun)
print(constr)
print(constr.dir)
print(rhs)
```

# First Delivery Constraints
```{r}
#Variables fuel purchases 1-6, corn purchases 1-6
#Constraints
# f1 f2 f3 f4 f5 f6 c1 c2 c3 c4 c5 c6
first_delivery_fuel <- c(1,0,0,0,0,0,0,0,0,0,0,0)
first_delivery_corn <- c(0,0,0,0,0,0,1,0,0,0,0,0)
constr <- matrix(c(first_delivery_fuel, first_delivery_corn), ncol=12, byrow=TRUE)
constr.dir <- c(">=", ">=")
rhs <- c(required_fuel[1], required_corn[1])

obj.fun <- c(fuel_costs, corn_costs)

#Preview problem definition
print(obj.fun)
print(constr)
print(constr.dir)
print(rhs)
```

# Run Model To Optimize Parameters
```{r}
#solving model
prod.sol <- lp("min", obj.fun , constr , constr.dir , rhs,compute.sens=TRUE)
prod.sol
```

# Verify Trivial Solution of First Shipment
```{r}
prod.sol$solution
```
```{r}
#Manually verifying trivial solution
2*420 + 1.5*780
```

## Additional Duality Check
```{r}
print("Minimized cost:")
prod.sol
#prod.sol$duals #includes duals of constraints and reduced costs of variables
```

## No More then triple fuel then corn
Example:
700 fuel and 200 corn violates the principle
1*700 -3*200 <= 0
700 - 600 <= 0
100 <= 0 #False; verifying that the principle has been violated
```{r}
january_constraint <- c(1,0,0,0,0,0,-3,0,0,0,0,0)
constr <- matrix(c(first_delivery_fuel, first_delivery_corn,
                   january_constraint
                   ), ncol=12, byrow=TRUE)
constr.dir <- c(">=",">=","<=")
rhs  <- c(required_fuel[1], required_corn[1], 0)

#Preview problem definition
print(obj.fun)
print(constr)
print(constr.dir)
print(rhs)

#resolving model
prod.sol <- lp("min", obj.fun , constr , constr.dir , rhs,compute.sens=TRUE)

print("Minimized cost:")
prod.sol
```
```{r}
prod.sol$solution
```

## Add Constraints For Q2 Shipment
Fuel from January will still be good for the Q2 shipment in April. Similarly, corn from January will also still be good. Thus the combined purchases from January and March must cover fuel and corn needs for both Q1 and Q2 shipments collectively.
```{r}
second_delivery_fuel <- c(1,1,0,0,0,0,0,0,0,0,0,0) 
second_delivery_corn <- c(0,0,0,0,0,0,1,1,0,0,0,0)
march_constraint <- c(0,1,0,0,0,0,0,-3,0,0,0,0)
constr <- matrix(c(first_delivery_fuel, first_delivery_corn,
                   january_constraint, march_constraint,
                   second_delivery_fuel, second_delivery_corn
                   ), ncol=12, byrow=TRUE)
constr.dir <- c(">=", ">=", "<=", "<=", ">=", ">=")
rhs <- c(required_fuel[1], required_corn[1],
         0,0,
         required_fuel[1]+required_fuel[2], required_corn[1]+required_corn[2]
         )
print(obj.fun)
print(constr)
print(constr.dir)
print(rhs)
```

# Reoptimize Solution
```{r}
#solving model
prod.sol <- lp("min", obj.fun , constr , constr.dir , rhs, compute.sens=TRUE)
```

# Preview Updated Results
```{r}
#Without march constraint: 1495    0    0    0    0    0  420  385    0    0    0    0
prod.sol$solution
print("Minimized cost:")
prod.sol
```

## Third Quarter Constraints
January, March, May, July
Only fuel purchased in May or July will be good for the Q3 shipment (fuel purchased in January or March will be expired in July when the Q3 delivery must be fulfilled). In the case of corn, only January's purchase will have exceeded its shelf life, so purchases from March, May and July are all acceptable. This constraint becomes a little tricky, as corn purchased in March may have been allocated to the Q2 shipment sent out in April.
```{r}
#JanFuel,MarchFuel,MayFuel,JulyFuel,SeptFuel,NovFuel,JanCorn,MarchCorn,MayCorn,JulyCorn,SeptCorn,NovCorn
third_delivery_fuel <- c(0,0,1,1,0,0,
                         0,0,0,0,0,0) 
third_delivery_corn_p1 <- c(0,0,0,0,0,0,
                            0,1,1,1,0,0)
third_delivery_corn_p2 <- c(0,0,0,0,0,0,
                            1,1,1,1,0,0)
may_constraint <- c(0,0,1,0,0,0,
                    0,0,-3,0,0,0)
july_constraint <- c(0,0,0,1,0,0,
                    0,0,0,-3,0,0)
constr <- matrix(c(first_delivery_fuel, first_delivery_corn,
                   january_constraint, march_constraint,
                   second_delivery_fuel, second_delivery_corn,
                   may_constraint, july_constraint,
                   third_delivery_fuel, third_delivery_corn_p1, third_delivery_corn_p2
                   ),
                 ncol=12, byrow=TRUE)
constr.dir <- c(">=", ">=",
                "<=", "<=",
                ">=", ">=",
                "<=", "<=",
                ">=", ">=", ">=")
rhs <- c(required_fuel[1], required_corn[1],
         0,0,
         required_fuel[1]+required_fuel[2], required_corn[1]+required_corn[2],
         0,0,
         required_fuel[3], required_corn[2]+required_corn[3], required_corn[1]+required_corn[2]+required_corn[3])
print(obj.fun)
print(constr)
print(constr.dir)
print(rhs)
```


# Reoptimize Solution
```{r}
#solving model
prod.sol <- lp("min", obj.fun , constr , constr.dir , rhs, compute.sens=TRUE)
```

# Preview Updated Results
```{r}
prod.sol$solution
print("Minimized cost:")
prod.sol
```

## Fourth Quarter Constraints

Quarter 4 delivery is set for early October. Therefore Fuel from July and September will be usable. Corn oil from May,July and September will be usable.
```{r}

fourth_delivery_fuel_p1 <- c(0,0,1,1,1,0,
                         0,0,0,0,0,0)
fourth_delivery_fuel_p2 <- c(0,0,0,1,1,0,
                         0,0,0,0,0,0) 
fourth_delivery_corn_p1 <- c(0,0,0,0,0,0,
                            0,0,1,1,1,0)
fourth_delivery_corn_p2 <- c(0,0,0,0,0,0,
                            1,1,1,1,1,0)
september_constraint <- c(0,0,0,0,1,0,
                    0,0,0,0,-3,0)



constr <- matrix(c(first_delivery_fuel, first_delivery_corn,
                   january_constraint, march_constraint,
                   second_delivery_fuel, second_delivery_corn,
                   may_constraint, july_constraint,
                   third_delivery_fuel, third_delivery_corn_p1, third_delivery_corn_p2,
                   fourth_delivery_fuel_p1, fourth_delivery_fuel_p2, fourth_delivery_corn_p1, fourth_delivery_corn_p2,
                   september_constraint
                   ),
                 ncol=12, byrow=TRUE)
constr.dir <- c(">=", ">=",
                "<=", "<=",
                ">=", ">=",
                "<=", "<=",
                ">=", ">=", ">=",
                ">=", ">=", ">=", ">=",
                "<=")
rhs <- c(required_fuel[1], required_corn[1],
         0,0,
         required_fuel[1]+required_fuel[2], required_corn[1]+required_corn[2],
         0,0,
         required_fuel[3], required_corn[2]+required_corn[3], required_corn[1]+required_corn[2]+required_corn[3],
         required_fuel[4], required_fuel[3]+required_fuel[4], required_corn[4],required_corn[1]+required_corn[2]+required_corn[3]+required_corn[4],
         0)
```

```{r}
print(obj.fun)
print(constr)
print(constr.dir)
print(rhs)
```

```{r}
prod.sol <- lp("min", obj.fun , constr , constr.dir , rhs, compute.sens=TRUE)
prod.sol
```

```{r}
prod.sol$solution
```



## Addendum

What if you get a 25% discount to any purchase over 1000 tons?

To do this we will need new variables f12-f62/c12-c62 to represent purchases over 1000 tons, while keeping the original f1-f6/c1-c6 variables for purchases under 1000 tons.
In order to specify the over 1000 constraint, we'll also need some binary variables, which will be 0 if we purchase under 1000 tons, or 1 if we purchase over 1000 tons.

In general, the constraints should generally translate naturally. For example, the first constraint is now:


```{r}
#Copied from above
constr <- matrix(c(first_delivery_fuel, first_delivery_corn,
                   january_constraint, march_constraint,
                   second_delivery_fuel, second_delivery_corn,
                   may_constraint, july_constraint,
                   third_delivery_fuel, third_delivery_corn_p1, third_delivery_corn_p2,
                   fourth_delivery_fuel_p1, fourth_delivery_fuel_p2, fourth_delivery_corn_p1, fourth_delivery_corn_p2,
                   september_constraint
                   ),
                 ncol=12, byrow=TRUE)
constr.dir <- c(">=", ">=",
                "<=", "<=",
                ">=", ">=",
                "<=", "<=",
                ">=", ">=", ">=",
                ">=", ">=", ">=", ">=",
                "<=")
rhs <- c(required_fuel[1], required_corn[1],
         0,0,
         required_fuel[1]+required_fuel[2], required_corn[1]+required_corn[2],
         0,0,
         required_fuel[3], required_corn[2]+required_corn[3], required_corn[1]+required_corn[2]+required_corn[3],
         required_fuel[4], required_fuel[3]+required_fuel[4], required_corn[4],required_corn[1]+required_corn[2]+required_corn[3]+required_corn[4],
         0)

#Constraints duplicated to allow for additional variables representing purchases over 1000
binary <- matrix(rep(0,dim(constr)[1]*12), ncol=12)
constr <- cbind(constr, constr, binary)
print(dim(constr))

#Now we need to add additional constraints specifying variables 1-12 are <= 1000 while variables 13-24 are >1000.
#All variables greater then or equal to 0. Binary variables less then or equal to 1.
for (i in 1:36){
  new <- rep(0,36)
  new[i] <- 1
  constr <- rbind(constr, new)
  constr.dir <- c(constr.dir, ">=")
  rhs <- c(rhs, 0)
  #Variables 12-24 (purchases over 1000 tons) must be over 1000 tons.
  if ((i>12)&&(i<=24)){
    new <- rep(0,36)
    new[i] <- 1 #Purchase over 1000 tons
    i2 <- i + 12
    new[i2] <- -1000 #Binary counterpart
    constr <- rbind(constr, new) #Ensure variables 12-24 are over 1000 (binary will be 1)
    constr.dir <- c(constr.dir, ">=")
    rhs <- c(rhs, 0)
    #Ensure binary variable is 1 when variables 12-24 > 0
    new <- rep(0,36)
    new[i] <- 1 #Purchase over 1000 tons
    i2 <- i + 12
    new[i2] <- -10000 #Arbitrary large number (we're ensuring binary variables aren't set to 0 when counterpart populated)
    constr <- rbind(constr, new) #Ensure variables 12-24 are over 1000 (binary will be 1)
    constr.dir <- c(constr.dir, "<=")
    rhs <- c(rhs, 0)
  }
  if (i > 24){
      new <- rep(0,36)
      new[i] <- 1
      constr <- rbind(constr, new)
      constr.dir <- c(constr.dir, "<=")
      rhs <- c(rhs, 1)
  }
}
print(dim(constr))
print(constr)
print(constr.dir)
print(rhs)
```


```{r}
#Finally, modify the objective function
obj.fun <- c(fuel_costs, corn_costs, .75*fuel_costs, .75*corn_costs, rep(0,12))
prod.sol <- lp("min", obj.fun , constr , constr.dir , rhs, compute.sens=TRUE, binary.vec=25:36)
prod.sol
```

```{r}
prod.sol$solution
```

## Discount Constraints

```{r}
obj.fun
```

## Just Discount Corn
```{r}
#Copied from above
constr <- matrix(c(first_delivery_fuel, first_delivery_corn,
                   january_constraint, march_constraint,
                   second_delivery_fuel, second_delivery_corn,
                   may_constraint, july_constraint,
                   third_delivery_fuel, third_delivery_corn_p1, third_delivery_corn_p2,
                   fourth_delivery_fuel_p1, fourth_delivery_fuel_p2, fourth_delivery_corn_p1, fourth_delivery_corn_p2,
                   september_constraint
                   ),
                 ncol=12, byrow=TRUE)
```

```{r}
#Copied from above
constr <- matrix(c(first_delivery_fuel, first_delivery_corn,
                   january_constraint, march_constraint,
                   second_delivery_fuel, second_delivery_corn,
                   may_constraint, july_constraint,
                   third_delivery_fuel, third_delivery_corn_p1, third_delivery_corn_p2,
                   fourth_delivery_fuel_p1, fourth_delivery_fuel_p2, fourth_delivery_corn_p1, fourth_delivery_corn_p2,
                   september_constraint
                   ),
                 ncol=12, byrow=TRUE)
constr.dir <- c(">=", ">=",
                "<=", "<=",
                ">=", ">=",
                "<=", "<=",
                ">=", ">=", ">=",
                ">=", ">=", ">=", ">=",
                "<=")
rhs <- c(required_fuel[1], required_corn[1],
         0,0,
         required_fuel[1]+required_fuel[2], required_corn[1]+required_corn[2],
         0,0,
         required_fuel[3], required_corn[2]+required_corn[3], required_corn[1]+required_corn[2]+required_corn[3],
         required_fuel[4], required_fuel[3]+required_fuel[4], required_corn[4],required_corn[1]+required_corn[2]+required_corn[3]+required_corn[4],
         0)

#Constraints duplicated to allow for additional variables representing purchases over 1000
binary <- matrix(rep(0,dim(constr)[1]*6), ncol=6)
constr <- cbind(constr, constr[,7:12], binary)
print(dim(constr))

#Now we need to add additional constraints specifying variables 1-12 are <= 1000 while variables 13-24 are >1000.
#All variables greater then or equal to 0. Binary variables less then or equal to 1.
for (i in 1:24){
  new <- rep(0,24)
  new[i] <- 1
  constr <- rbind(constr, new)
  constr.dir <- c(constr.dir, ">=")
  rhs <- c(rhs, 0)
  #Variables 12-18 (purchases over 1000 tons) must be over 1000 tons.
  if ((i>12)&&(i<=18)){
    new <- rep(0,24)
    new[i] <- 1 #Purchase over 1000 tons
    i2 <- i + 6
    new[i2] <- -1000 #Binary counterpart
    constr <- rbind(constr, new) #Ensure variables 12-24 are over 1000 (binary will be 1)
    constr.dir <- c(constr.dir, ">=")
    rhs <- c(rhs, 0)
    #Ensure binary variable is 1 when variables 12-24 > 0
    new <- rep(0,24)
    new[i] <- 1 #Purchase over 1000 tons
    i2 <- i + 6
    new[i2] <- -10000 #Arbitrary large number (we're ensuring binary variables aren't set to 0 when counterpart populated)
    constr <- rbind(constr, new) #Ensure variables 12-24 are over 1000 (binary will be 1)
    constr.dir <- c(constr.dir, "<=")
    rhs <- c(rhs, 0)
  }
  if (i > 18){
      new <- rep(0,24)
      new[i] <- 1
      constr <- rbind(constr, new)
      constr.dir <- c(constr.dir, "<=")
      rhs <- c(rhs, 1)
  }
}
print(dim(constr))
print(constr)
print(constr.dir)
print(rhs)
```

```{r}
#Finally, modify the objective function
obj.fun <- c(fuel_costs, corn_costs, .75*corn_costs, rep(0,6))
prod.sol <- lp("min", obj.fun , constr , constr.dir , rhs, compute.sens=TRUE, binary.vec=19:24)
prod.sol
```
