---
title: "Aviation Bio-Fuel Stock Sourcing Problem"
output: html_notebook
---

# Problem Statement

A company that produces aircraft biofuel is planning a new product
called FC (Fuel-Corn). The table below shows the total quarterly demand in
tons (t) for the coming years as communicated by their customers.

FC demand (T)

| Q1             | Q2           | Q3           | Q4          |
| :------------- | :----------: | -----------: |-----------: |
| 1,200    | 1,100    | 1,300   |  1,000 |

Fuel corn is made of a combination of Fuel and Corn. The supply schedule dictates that we purchase these raw materials on a bimonthly schedule. The costs per ton of Fuel and Corn for
every two month period in the coming year is listed below.

|           | B1        | B2      |    B3    | B4   |  B5  |  B6  |
| :-------- | :-------: | :------:| :------: | :---:|:----:| ----:|
|Fuel (\$/t) | 2         |2.5      |   2      |  1   |  1.5 |  3   |
|Corn (\$/t) | 1.5       | 1       |   2      |  1   |  2   |   2.5|


FC composition is obtained by mixing 35% of Fuel and 65% of Corn.
The life of Fuel is of four consecutive months and the life of Corn, six
(i.e., if we buy Fuel in early January, we cannot use it in early May).
We just buy Fuel and Corn at the beginning of each two-month period
and make the deliveries of FC at the beginning of each quarter. For
simplicity, we assume that one can buy, mix and sell the same day.
In addition, the plant manager has told us that in any two-month period, we cannot buy more Fuel than triple of Corn.

Develop a model to minimize the cost of production.

# Decision Variables
bf[1:6] buying period for fuel (life is 4 months)
bc[1:6] buying period for corn (life is 6 months)

bf[1] can go to Q1 and Q2 
bf11 + bf12

bf[2] can go to Q2
bf22

bf[3] can go to Q3
bf33

bf[4] can go to Q3 and Q4
bf43 + bf44

bf[5] can go to Q4
bf54

bc[1] can go to Q1 and Q2
bc11 + bc12

bc[2] can go to Q2 and Q3
bc22 + bc23

bc[3] can go to Q3 and Q4
bc33 + bc34

bc[4] can go to Q3 and Q4
bc43 + bc44

bc[5] can go to Q4
bc54

Q1 is 1
Q2 is 4
Q3 is 7
Q4 is 10

```{r}

m <- matrix(c(1,1,4,6,2,3,6,8,3,5,8,10,4,7,10,12,5,9,12,14,6,11,14,16), ncol = 4, byrow = TRUE)
colnames(m) <- c("B", "Start", "Fuel End", "Corn End")
print(m)


```

# Contraints
bf11 + bf12 - 3 (bc11 + bc12) <= 0
bf22 - 3 (bc22 + bc23) <= 0
bf33 - 3 (bc33 + bc34) <= 0
bf43 + bf44 - 3 (bc43 + bc44) <= 0
bf54 - 3 (bc54) <= 0

# Q1
bf11 = 420
bc11 = 780

# Q2
bf12 + bf22 = 385
bc12 + bc22 = 715

# Q3
bf33 + bf43 = 455
bc23 + bc33 + bc43 = 845

#Q4
bf44 + bf54 = 350
bc34 + bc44 + bc54 = 650

# objective function
1 bf11
2 bf12
3 bf22
4 bf33
5 bf43
6 bf44
7 bf54
8 bc11
9 bc12
10 bc22
11 bc23
12 bc33
13 bc34
14 bc43
15 bc44
16 bc54

17-32: Discount Counterparts
32-48: Binary Counterparts

2 (bf11 + bf12) + 2.5 bf22 + 2 bf33 + 1 (bf43 + bf44) + 1.5 (bf54) +
1.5 (bc11 + bc12) + 1(bc22 + bc23) + 2 (bc33 + bc34) + 1 (bc43 + bc44) + 2 (bc54)

```{r}
library(lpSolveAPI)

model <- make.lp(0, 16*3)

set.bounds(model, lower = rep(0,16*3))

standard_rates <- c(2,2,
                   2.5,
                   2,
                   1,1,
                   1.5,
                   
                   1.5,1.5,
                   1,1,
                   2,2,
                   1,1,
                   2)
set.objfn(model, c(standard_rates, .75*standard_rates, rep(0,16)))

add.constraint(model, rep(c(1,1,-3,-3),2),"<=", 0, c(c(1,2,8,9),c(1,2,8,9)+16))
add.constraint(model, rep(c(1,-3,-3),2),"<=", 0, c(c(3,10,11),c(3,10,11)+16))
add.constraint(model, rep(c(1,-3,-3),2),"<=", 0, c(c(4,12,13),c(4,12,13)+16))
add.constraint(model, rep(c(1,1,-3,-3),2),"<=", 0, c(c(5,6,14,15),c(5,6,14,15)+16))
add.constraint(model, rep(c(1,-3),2),"<=", 0, c(c(7,16),c(7,16)+16))

# Q1 constraints
add.constraint(model, rep(1,2),">=", 420, c(1,1+16))
add.constraint(model, rep(1,2),">=", 780, c(8,8+16))

# Q2 constraints
add.constraint(model, rep(1,4),">=", 385, c(c(2,3),c(2,3)+16))
add.constraint(model, rep(1,4),">=", 715, c(c(9,10),c(9,10)+16))

# Q3 constraints
add.constraint(model, rep(1,4),">=", 455, c(c(4,5),c(4,5)+16))
add.constraint(model, rep(1,6),">=", 845, c(c(11,12,13),c(11,12,13)+16))

# Q4 constraints
add.constraint(model, rep(1,4),">=", 350, c(c(6,7),c(6,7)+16))
add.constraint(model, rep(1,6),">=", 650, c(c(14,15,16),c(14,15,16)+16))


solve(model)

print(sum(get.variables(model)[1:2]))
print(sum(get.variables(model)[3]))
print(sum(get.variables(model)[4]))
print(sum(get.variables(model)[5:6]))
print(sum(get.variables(model)[7]))

print(sum(get.variables(model)[8:9]))
print(sum(get.variables(model)[10:11]))
print(sum(get.variables(model)[12:13]))
print(sum(get.variables(model)[14:15]))
print(sum(get.variables(model)[16]))

print(sum(get.variables(model)[c(1,8)]))
print(sum(get.variables(model)[c(2,3,9,10)]))
print(sum(get.variables(model)[c(4,5,11,12,13)]))
print(sum(get.variables(model)[c(6,7,14,15,16)]))

get.objective(model)

```